---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# iosoi (*I*dentifier *O*f *S*olution *O*f *I*nterest)

<!-- badges: start -->
<!-- badges: end -->

The goal of *iosoi* is to provide an [R](https://www.r-project.org) 
implementation of the multi-criteria decision making approach proposed in:

- Torres, M., Pelta, D. A., Lamata, M. T., & Yager, R. R. (2021). 
An approach to identify solutions of interest from multi and many-objective 
optimization problems. *Neural Computing & Applications, 33*(7), 2471–2481. 
[doi:10.1007/s00521-020-05140-x](https://doi.org/10.1007/s00521-020-05140-x)

It also includes alternative approaches for assessing the solutions from a
geometric perspective. Formally, iosoi helps to solve the following MCDM problem:

Given a set of m solutions (alternatives) with evaluations on n criteria, 
and an order of preference among these criteria, **order the solutions according 
to their overall scores**. These overall scores are obtained through the linear 
combination of the evaluations and the weights of the criteria.

## Installation

You can install the development version of *iosoi* from 
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("pnovoa/iosoi")
```

## Example

Consider a decision problem in which the objective is to identify a set of 
solutions of interest based on their overall performance over a given set of 
criteria. As input data we have a set of $m$ solutions (alternatives) and their 
individual evaluations on each of the $n$ criteria. Intuitively, this 
information can be arranged in the form of a matrix $E$ whose entries $e_{ij}$ 
correspond to the numerical evaluation that solution i has in criterion j. 
Additionally, we have an order of preference of the criteria by the decision 
maker. Without loss of generality, it will be assumed that this order is 
decreasing as a function of the order of the criteria in the matrix $E$. 
That is, let $\lambda_j$ the $j-th$ criterion, then 
$\lambda_1 \succeq \lambda_2 ... \succeq \lambda_n$. Numerically, this 
information can be translated into variables representing non-negative 
weights associated with the criteria and with the additional condition of 
adding up to 1.

The following example is based on a problem of 5 solutions, 3 criteria and 
evaluation matrix:

$$ E_{5 \times 3} =
\left(\begin{array}{cc} 
5 & 3 & 2 \\
2 & 1 & 3 \\
3 & 1 & 4 \\
4 & 5 & 3 \\
4 & 4 & 2 \\
\end{array}\right)
$$ 

The preference order provided by the decision-maker is (as we mentioned before)
in decreasing order regarding the order followed by the criteria in matrix 
$E_{ij}$.

Using the `iosoi` package, one can perform the steps of the Torres et al. (2021) 
approach to obtain global evaluations of the solutions and be able to select 
those of greatest interest to the decision maker. In what follows we will perform
this task from both, the possibility approach by Torres et al. (2021) and a
geometric approach based on the notion of the volume. In the case of possibility
approach we will rely on the *neutral* attitude for computing the superiority
degree of the solutions against the reference one. For both approach we set
a threshold of $0$, which means that only those solutions with their 
corresponding assessments greater than $0$ will appear in the output.

```{r example}
library(iosoi)
# Evaluation matrix
E <- matrix(c(5, 2, 3, 4, 4, 3, 1, 1, 5, 4, 2, 3, 4, 3, 2), byrow = TRUE, nrow = 5)

# Identifying SOIs from the possibility approach of Torres et al (2021).
E %>% poss_identify_sois(by = "neutral", threshold = 0.0)

# Identifying SOIs from the geometric approach based on the volume
E %>% geom_identify_sois(by = "poss_volume", threshold = 0.0)
```
As noted, both outputs contain not only the original evaluation matrix, but also
other columns related to intermediate steps of the method of Torres et al (2021).
For instance, those columns with prefix `VE_` corresponds to the linear 
scoring of each solution at the extreme points of the feasible region induced by
the preference order of the criteria. Columns `LB` and `UB` contain the lower 
and upper bounds of the values obtained in the `VE_` columns. The Boolean column
`REF` identify (with 1) the reference solution, that is, the one used to assess 
the rest of solutions. Finally, the last column (ie. `neutral` or `poss_volume`) 
contains the overall assessment of the solutions. Note that only those with values
greater than $0$ were included. In the first case, only solutions `S1`, 
`S2`, `S4` and `S5` were selected, while in the second case all were included.

To understand why solution S3 is left out in the approach of Torres et al. 
(2021) `iosoi` allows to plot the intervals (ranges of possible scores) of each 
solution. That is, in order to visually check the degree of overlapping of 
the solutions in relation to the reference one.  
The code then performs the analysis again but relaxes the threshold to $-1$ 
to include all solutions in the final output. In this way the intervals of 
all the solutions will be plotted.

```{r plots}
E %>% 
  poss_identify_sois(by = "neutral", threshold = -1.0) %>%
  plot_intervals()
```

As expected, `S3` is excluded because its interval is clearly far to the left 
relative to `S2` (the reference solution highlighted with interval in red). 
The rest, and especially `S1` have certain levels of overlap with `S2`. 

We can also plot any computed assessment indicator through a bar plot. For 
example, if we consider `poss_volume`, the following code show how to plot
this indicator.

```{r plots2}
E %>% 
  geom_identify_sois(by = "poss_volume", threshold = 0.0) %>%
  plot_assessment(by = "poss_volume",
                  plot_title = "Normalized volume",
                  ylabel = "Normalized volume"
                  )
```




